---
output: hugodown::hugo_document

slug: three-new-tidymodels-packages
title: "Three New tidymodels Packages"
date: 2025-09-03
author: Max Kuhn, Frances Lin, Simon Couch
description: >
    A 2-3 sentence description of the post that appears on the articles page.
    This can be omitted if it would just recapitulate the title.

photo:
  url: https://unsplash.com/photos/three-french-macaroons-on-plate-71jYZb6Ag7M
  author: Keila Hötzel

# one of: "deep-dive", "learn", "package", "programming", "roundup", or "other"
categories: [package] 
tags: [tidymodels,postprocessing,feature-selection]
---

<!--
TODO:
* [ ] Look over / edit the post's title in the yaml
* [ ] Edit (or delete) the description; note this appears in the Twitter card
* [ ] Pick category and tags (see existing with `hugodown::tidy_show_meta()`)
* [ ] Find photo & update yaml metadata
* [ ] Create `thumbnail-sq.jpg`; height and width should be equal
* [ ] Create `thumbnail-wd.jpg`; width should be >5x height
* [ ] `hugodown::use_tidy_thumbnails()`
* [ ] Add intro sentence, e.g. the standard tagline for the package
* [ ] `usethis::use_tidy_thanks()`
-->

We're very chuffed to announce the release of _three_ new modeling packages: filtro, important, and tailor. 

You can install them from CRAN with:

```{r, eval = FALSE}
install.packages(c("filtro", "important", "tailor"))
```

This blog post will describe each. 

## filtro

Feature selection is an important step in building robust machine learning models. By keeping only the most relevant predictors, we can reduce overfitting, improve model interpretability, and speed up computation.

filtro is tidy tools designed specifically for filter-based supervised feature selection. filtro makes it easy to score, rank, and select features using a wide range of statistical and model-based metrics. These include: p-values, correlation, random forest feature importance, information gain, and more.

With filtro, we can quickly rank the variables and select either the top proportion or the top number features that best contribute to our model. 

First, we need to load the required packages:

```{r}
library(filtro)
library(desirability2)
library(dplyr)
library(modeldata)
```

The modeldata package contains a data set used to predict housing sale price. It has 73 predictor columns and a numeric variable Sale_Price (the outcome). Since the outcome are right-skewed, we apply a log (base 10) transformation.

```{r}
ames_subset <- modeldata::ames |>
  # Use a subset of data for demonstration
  dplyr::select(
    Sale_Price,
    MS_SubClass,
    MS_Zoning,
    Lot_Frontage,
    Lot_Area,
    Street
  )
ames_subset <- ames_subset |>
  dplyr::mutate(Sale_Price = log10(Sale_Price))
```

To apply the correlation filter, we first create a score class object to define the scoring method, and then use the `fit()` method with the standard formula to compute the scores. The data frame of results can be accessed via object@results.

```{r}
# Pearson correlation
ames_cor_pearson_res <-
  score_cor_pearson |>
  fit(Sale_Price ~ ., data = ames_subset)
ames_cor_pearson_res@results
```

Because the outcome is numeric, any predictor that is not a factor will result in an NA. In case where NA is produced, a safe value can be used to retain the predictor, and can be accessed via object@fallback_value.

We can also apply the random forest filter: 

```{r}
# Forest importance
ames_imp_rf_reg_res <-
  score_imp_rf |>
  fit(Sale_Price ~ ., data = ames_subset, seed = 42)
ames_imp_rf_reg_res@results
```

Similarly, we can apply the entropy-based filter: 

```{r}
# Information gain
ames_info_gain_reg_res <-
  score_info_gain |>
  fit(Sale_Price ~ ., data = ames_subset)
ames_info_gain_reg_res@results
```

Sometimes, we may want to combine multiple scores into one unified score. filtro also supports multiparameter optimization, powered by the desirability2 package.

To do this, we first collect the individual score objects into a list.  

```{r}
# Create a list
class_score_list <- list(
  ames_cor_pearson_res,
  ames_imp_rf_reg_res,
  ames_info_gain_reg_res
)
```

Then, we prepare the results by filling in safe values for NA, as well as removing the outcome column.

```{r}
# Fill safe values
ames_scores_results <- class_score_list |>
  fill_safe_values() |>
  # Remove outcome
  dplyr::select(-outcome)
ames_scores_results
```

Now, we can perform single- or multi-parameter optimization using desirability functions. The `show_best_desirability_prop()` method allows us to rank the scores according to some verbs, such as `maximize`, `minimize`, `target`, and `constrain`) and their associated criteria (e.g., `prop_terms = `). 

```{r}
# Single and multi-parameter optimization using desirability functions
# Optimize correlation alone
ames_scores_results |>
  show_best_desirability_prop(
    maximize(cor_pearson, low = 0, high = 1)
    # prop_terms defaults to 1 
  )
```

Here, the `.d_max_cor_pearson` column represents the desirability of the correlation score for each predictor, with $1.0$ being most desirable and $0.0$ unacceptable. In this example, predictors `MS_SubClass`, `MS_Zoning`, and `Street` achieve the maximum desirability of $1$ for correlation, while `Lot_Area` and `Lot_Frontage` are lower. The `.d_overall` column summarizes the overall desirability using the geometric mean. Since we are optimizing only a single parameter in this example, the `.d_max_cor_pearson` and `.d_overall` columns are identical. 

For more examples: 

```{r}
# Single and multi-parameter optimization using desirability functions
# Optimize correlation and forest importance
ames_scores_results |>
  show_best_desirability_prop(
    maximize(cor_pearson, low = 0, high = 1),
    maximize(imp_rf)
  )

# Optimize correlation, forest importance and information gain
ames_scores_results |>
  show_best_desirability_prop(
    maximize(cor_pearson, low = 0, high = 1),
    maximize(imp_rf),
    maximize(infogain)
  )

# Same as above, but retain only a proportion of predictors
ames_scores_results |>
  show_best_desirability_prop(
    maximize(cor_pearson, low = 0, high = 1),
    maximize(imp_rf),
    maximize(infogain),
    prop_terms = 0.2
  )

# Optimize toward a target
ames_scores_results |>
  show_best_desirability_prop(
    target(cor_pearson, low = 0.2, target = 0.255, high = 0.9)
  )

# Optimize with box constraints
ames_scores_results |>
  show_best_desirability_prop(
    constrain(cor_pearson, low = 0.2, high = 1)
  )
```

Currently, filtro implements a total of six filters. For more details and examples, [please visit the package site](https://filtro.tidymodels.org). Developers can also [check out the post on how to create a custom score object](https://www.tidymodels.org/learn/develop/filtro/). 

## important

## tailor

## Acknowledgements

